---
author: "Yuhaniz Aly"
title: "Colombia Health Loss"
output: html_document
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
source("analysis.R")
library(dplyr)
library(knitr)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

# Columbia Health Loss

## Summary 


## Deaths
One metric of disease burden is the number of *deaths*. Below is a barchart of 
the top 10 diseases that cause the most health loss in Colombia:

```{r plot1, echo = FALSE, warning=FALSE, fig.width = 12, fig.height = 6}
top_10_deaths_plot
```

## YLLs
Another metric of disease burden is *Years of Life Lost (YLLs)*. This measures 
the years of potential life lost due to premature deaths. Below is a bar chart
of the top 10 diseases as measured by
YLLs. 
```{r plot2, echo = FALSE, warning=FALSE, fig.width = 12, fig.height = 6}
top_10_ylls_plot
```


## YLDs
Another metric of disease burden is *Years of Lived with Disability (YLDs)*.
This measures the years of potential life lost
due to living with a disability. Below is a bar chart of the top 10 diseases as measured by YLDs. 
```{r plot3, echo = FALSE, warning=FALSE, fig.width = 12, fig.height = 6}
top_10_ylds_plot
```


## DALYs
Lastly, another metric of disease burden is
*Disability Adjusted Life Years (DALYs)*. This measures the overall disease 
burden, expressed the years of potential life lost due to ill-health,
disability, or premature death. DALYs combines mortality and morbidity into one
metric. Below is a bar chart of the top 10 diseases as measured by DALYs. 
```{r plot4, echo = FALSE, warning=FALSE, fig.width = 12, fig.height = 6}
top_10_dalys_plot
```

## Comparison of  Asthma
Below is a barchart of different health metrics comparing asthma.
```{r plot5, echo = FALSE, warning=FALSE, fig.width = 12, fig.height = 6}
asthma_comparison
```
Based on the barchart, asthma has the most ylds and the least deaths. This means
asthma is not a deadly health condition as not many people die from this cause.
This is also shown for the ylls comparison because not many lifetime years are 
lost from asthma. In contrast, a higher percentage of people end up living with 
asthma, which explains the large percentage of ylds. Lastly, the DALYs tells us 
that the overall burden for asthma is about 1.7%.